Sound Automated Synthesis of Digital Stabilizing Controllers for Continuous Plants

### 2016-10-13 paper/paper.tex
wordings

%------------------------------------
\subsection{Sampling and Holding in Closed-loop Systems}
\label{ssec:SandH}
%------------------------------------

The digital controllers synthesised using the algorithm we present in this
paper are typically used in closed loops with continuous (and physical)
plants.  Thus, the proposed approach has to consider dynamics of continuous
parts of the system (i.e., the plant) and the discrete parts (i.e., the
digital controller).  To obtain a joint model, the discretise the continuous
plant.  We only consider transfer function models, and require a $z$-domain
transfer function that captures all aspects of the continuous plant (with a
Laplace domain transfer function $G(s)$) and the sampling and hold process.

Note that $G(z)$ is a discrete model of a real-domain (and continuous)
plant, i.e., the continuous model of the plant must be discretised to obtain
the discrete equivalent coefficients via ZOH discretisation.

@@

In this study, the methodology used for verifying closed-loop digital
control system is based on the Digital-System Verifier
(DSVerifier)~\cite{IsmailBCFF15}, which checks the stability of control
systems considering finite-word length (FWL) effects in the digital
controller and uncertainty parameters in the plant model (plant
intervals)~\cite{Bessa16}.  Let $C(z)$ be a digital controller and $G(z)$ be
a plant model given as
%
\begin{align}
\small
\label{controller_plant_tf}
C(z)&=\frac{C_n(z)}{C_d{z}}=\frac{\beta_{0}+\beta_{1}z^{-1}+...+\beta_{M_C}z^{-M_C}}{\alpha_{0}+\alpha_{1}z^{-1}+...+\alpha_{N_C}z^{-N_C}}, \\
%\label{plant_tf}
G(z)&=\frac{G_n(z)}{G_d(z)}=\frac{b_{0}+b_{1}z^{-1}+...+b_{M_G}z^{-M_G}}{a_{0}+a_{1}z^{-1}+...+a_{N_G}z^{-N_G}}.
\end{align}
%
\noindent where $\vec{\beta}$ and $\vec{\alpha}$ are vectors containing the
controller's coefficients; similarly $\vec{b}$ and $\vec{a}$ encompass the
plant's coefficients; and $N_{(\cdot)}$ and $M_{(\cdot)}$ denote the number
of poles and zeros for the corresponding block.

Uncertainties in $G(z)$ may appear due to uncertainties in $G(s)$
(\emph{i.e.} $\Delta_p{G}_n(s),\Delta_p{G}_d(s)$), errors in the numeric
calculations, or roundoffs in the quantisation.  These uncertainties may be
parametrically expressed by additive terms, resulting in a uncertain model
$\hat{G}(z)$, such that:
%
\begin{equation}
\hat{G}(z)=\frac{G_n(z)+\Delta G_n(z)}{G_d(z)+\Delta G_d(z)},
\end{equation}
%
which will be represented by the following transfer function:
%
\begin{equation}
\hat{G}(z)=\frac{\hat{b}_{0}+\hat{b}_{1}z^{-1}+...+\hat{b}_{M_G}z^{-M_G}}{\hat{a}_{0}+\hat{a}_{1}z^{-1}+...+\hat{a}_{N_G}z^{-N_G}} : N_G \geq M_G .
\end{equation}
%
Due to the nature of the methods used for stability check in this paper, we
require that the parametric errors in the plant have the same polynomial
order as the plant itself (other errors are ensured to fulfil this property
in our implementation).

In order to simplify the notations in this paper, we will use a notation
based on the coefficients of the polynomials.  Let $\mathcal{P}^{N}$ be a
polynomial domain of order $N$ Let $P \in \mathcal{P}^{M,N}$ be a rational
polynomial $\frac{P_n}{P_d} : P_n \in \mathcal{P}^{M} \wedge P_d \in
\mathcal{P}^{N}$.  There is a vector of coefficients
%
\begin{equation}
\vec{P} \in \mathbb{R}^{N+M+2}=[n_{0}\ n_{1}\ \hdots \ n_{M}\ d_{0}\ d_{1}\ \hdots\ d_{N}\ ]^T
\label{eq:coefficients}
\end{equation}
%
with uncertainty
%
\begin{equation}
\Delta{\vec{P}}=[\Delta{n}_{0}\ \Delta{n}_{1}\ \hdots \ \Delta{n}_{M}\ \Delta{d}_{0}\ \Delta{d}_{1}\ \hdots\ \Delta{d}_{N}\ ]^T
\label{eq:coefficients}
\end{equation}
%
Then $\vec{\hat{G}}=\vec{G}+\Delta{\vec{G}}$.  In the following we may use
either notation to represent the functions $G(z)$, $C(z)$, and their
respective uncertainties.\\

@@

A typical digital control system system, with a continuous plant and a discrete controller,
is shown in Figure~\ref{fig:sampledsystem}, where the DAC and ADC converters
introduce quantisation noises $\nu_{1}(z)$ and $\nu_{2}(z)$, $G(s)$ is the
continuous-time plant model with parametric additive uncertainty $\Delta_p
\vec{G}_s$ , $R(z)$ is the reference signal, $U(z)$ is the control signal,
and $\hat{Y}(z)$ is the output signal affected by the noises and
uncertainties in the closed-loop system.

@@

The ADC and DAC may be abstracted changing the hybrid system in
figure~\ref{fig:sampledsystem} to the digital system in
figure~\ref{fig:digsystem1} where the parametric uncertainties are
represented by $\Delta_p \vec{G}$
%
%=\mathcal{Z}\left\lbrace{G_{P}(s)}\right\rbrace$
%
and the entire effect of $\nu_{1}$ and $\nu_{2}$ in the output $Y(z)$ is
represented as an additive (output) measurement noise $\nu(z)$, thus
%$\hat{Y}(z)=Y(z)+\nu(z)$.  The measurement noise $\nu(k)$ is enough for
%represent both quantisation noises ($\nu_{1}$ and $\nu_{2}$) what can be
%obtained reorganizing the equations~\ref{eq:outputfunctions}:
%%
%\begin{equation}
%\label{eq:outputfunctions2}
%\hat{Y}(z)=H_{1}R(z)+H_{1}\nu_{1}(z)+H_{2}(z)\nu_{2}(z),
%\end{equation}
%%
%where $H_{1}R(z)$ is exactly equal to $Y(z)$ without quantisation noise effects, thus:
%%
%\begin{equation}
%\nu(z)=H_{1}\nu_{1}(z)+H_{2}(z)\nu_{2}(z).
%\end{equation}

@@

%
In figure~\ref{fig:digsystem1}, two sources of uncertainties are considered: parametric uncertainties related to modelling errors, and uncertainties introduced by quantisations in sampling and holding process, {\it i.e.} ADC and DAC conversions. Assuming linearity and independence between these sources, the uncertain model may be rewritten as follows:
\begin{equation}
\label{eq:complete_unc_model}
\vec{\hat{G}}=\vec{G}+\Delta_p \vec{G}+\Delta_q \vec{G},
\end{equation}
%
where $\Delta_p \vec{G}$ is the term that corresponds to the parametric
uncertainties and $\Delta_q \vec{G}$ is the term that reflects the
quantisation effects on the plant model.

@@

Since the controller is implemented using a FWL, $C(z)$ also suffers
disturbances from the FWL effects, with roundoffs in coefficients that may
change the closed-loop poles and zeros position, and consequently affect its
stability, as shown in~\cite{Bessa16}.

Let $\hat{C}(z)$ be the transfer function of the digital controller
represented using this FWL with integer size $I$ and fractional size $F$. 
We shall denote the FWL domain of coefficients as $\mathbb{R}\langle I,F
\rangle$
%
%Since the controller is implemented in the $\mathbb{R}\langle I,F \rangle$ 
%domain, {\it i.e.} a Fixed-Point Finite Word Length with integer size $I$ and fractional
%size $F$, we shall denote this domain as $\mathbb{R}\langle I,F \rangle$ 
and assume that there exists a function $FWL[\cdot]:\mathcal{P}^{n}\rightarrow \mathcal{P}^{n}\langle I,F \rangle$, which applies the FWL effects to a polynomial. $\mathcal{P}^{n}$ is the space of polynomials of $n$-th order and $\mathcal{P}^{n}\langle I,F \rangle$ is is the space of polynomials with coefficients in $\mathbb{R}\langle I,F \rangle$.
%, \red{[explain index n! why do you need to work on a Euclidean domain?]}. 
%
Thus the perturbed controller model $\hat{C}(z)$ may be computed from the
original model $C(z)=\frac{N_{C}(z)}{D_{C}(z)}$ using the following
expression:
%
\begin{equation}
\hat{C}(z)=\frac{FWL[N_{C}(z)]}{FWL[D_{C}(z)]}.
\end{equation}
%
In the case of a synthesised controller, because the synthesis is performed
using FWL, $\hat{C}(z) \equiv C(z)$.

>>>>

%------------------------------------
\subsection{Discretising the Plant}
\label{ssec:SandH}
%------------------------------------

The digital controllers synthesised using the algorithm we present in this
paper are typically used in closed loops with continuous (and physical)
plants.  Thus, the proposed approach has to consider dynamics of continuous
parts of the system (i.e., the plant) and the discrete parts (i.e., the
digital controller).  To obtain a joint model, the discretise the continuous
plant. We briefly recall the standard foundations of this step.

We only consider transfer function models, and require a $z$-domain
transfer function that captures all aspects of the continuous plant (with a
Laplace domain transfer function $G(s)$) and the sampling and hold process.
Note that $G(z)$ is a discrete model of a real-domain (and continuous)
plant, i.e., the continuous model of the plant must be discretised to obtain
the discrete equivalent coefficients via ZOH discretisation.

@@

CEGIS-based control synthesis requires a formal verifier to check whether a
candidate controller meets the requirements when combined with the plant. 
We use the Digital-System Verifier (DSVerifier)~\cite{IsmailBCFF15} for this
purpose.  It checks the stability of closed-loop control systems and
considers finite-word length (FWL) effects in the digital controller and
uncertainty parameters in the plant model (plant intervals)~\cite{Bessa16}. 
Let $C(z)$ be a digital controller and $G(z)$ be a plant model given as
%
\begin{align}
\small
\label{controller_plant_tf}
C(z)&=\frac{C_n(z)}{C_d{z}}=\frac{\beta_{0}+\beta_{1}z^{-1}+...+\beta_{M_C}z^{-M_C}}{\alpha_{0}+\alpha_{1}z^{-1}+...+\alpha_{N_C}z^{-N_C}}, \\
%\label{plant_tf}
G(z)&=\frac{G_n(z)}{G_d(z)}=\frac{b_{0}+b_{1}z^{-1}+...+b_{M_G}z^{-M_G}}{a_{0}+a_{1}z^{-1}+...+a_{N_G}z^{-N_G}}.
\end{align}
%
\noindent where $\vec{\beta}$ and $\vec{\alpha}$ are vectors containing the
controller's coefficients; similarly $\vec{b}$ and $\vec{a}$ encompass the
plant's coefficients; and $N_{(\cdot)}$ and $M_{(\cdot)}$ denote the number
of poles and zeros for the corresponding block.

Uncertainties in $G(z)$ may appear due to uncertainties in $G(s)$
(\emph{i.e.} $\Delta_p{G}_n(s),\Delta_p{G}_d(s)$), errors in the numeric
calculations, or roundoffs in the quantisation.  These uncertainties may be
parametrically expressed by additive terms, resulting in a uncertain model
$\hat{G}(z)$, such that:
%
\begin{equation}
\hat{G}(z)=\frac{G_n(z)+\Delta G_n(z)}{G_d(z)+\Delta G_d(z)},
\end{equation}
%
which will be represented by the following transfer function:
%
\begin{equation}
\hat{G}(z)=\frac{\hat{b}_{0}+\hat{b}_{1}z^{-1}+...+\hat{b}_{M_G}z^{-M_G}}{\hat{a}_{0}+\hat{a}_{1}z^{-1}+...+\hat{a}_{N_G}z^{-N_G}} : N_G \geq M_G .
\end{equation}
%
Due to the nature of the methods we use for the stability check, we require
that the parametric errors in the plant have the same polynomial order as
the plant itself (other errors are ensured to fulfil this property in our
implementation).

In order to simplify presentation, we will use notation based on the
coefficients of the polynomials.  Let $\mathcal{P}^{N}$ be a polynomial
domain of order $N$ Let $P \in \mathcal{P}^{M,N}$ be a rational polynomial
$\frac{P_n}{P_d} : P_n \in \mathcal{P}^{M} \wedge P_d \in \mathcal{P}^{N}$. 
There is a vector of coefficients
%
\begin{equation}
\vec{P} \in \mathbb{R}^{N+M+2}=[n_{0}\ n_{1}\ \hdots \ n_{M}\ d_{0}\ d_{1}\ \hdots\ d_{N}\ ]^T
\label{eq:coefficients}
\end{equation}
%
with uncertainty
%
\begin{equation}
\Delta{\vec{P}}=[\Delta{n}_{0}\ \Delta{n}_{1}\ \hdots \ \Delta{n}_{M}\ \Delta{d}_{0}\ \Delta{d}_{1}\ \hdots\ \Delta{d}_{N}\ ]^T \;.
\label{eq:coefficients}
\end{equation}
%
Then $\vec{\hat{G}}=\vec{G}+\Delta{\vec{G}}$.  In the following we may use
either notation to represent the functions $G(z)$, $C(z)$, and their
respective uncertainties.\\

@@

A typical digital control system system with continuous plant and discrete
controller is illustrated in Figure~\ref{fig:sampledsystem}.  The DAC and
ADC converters introduce quantisation noises $\nu_{1}(z)$ and $\nu_{2}(z)$,
$G(s)$ is the continuous-time plant model with parametric additive
uncertainty $\Delta_p \vec{G}_s$, $R(z)$ is the reference signal, $U(z)$ is
the control signal, and $\hat{Y}(z)$ is the output signal affected by the
noises and uncertainties in the closed-loop system.

@@

The ADC and DAC may be abstracted by transforming the hybrid system in
Figure~\ref{fig:sampledsystem} into the digital system in
Figure~\ref{fig:digsystem1}, where the parametric uncertainties are
represented by~$\Delta_p \vec{G}$
%
%=\mathcal{Z}\left\lbrace{G_{P}(s)}\right\rbrace$
%
and the entire effect of $\nu_{1}$ and $\nu_{2}$ in the output $Y(z)$ is
represented as an additive (output) measurement noise $\nu(z)$.
%, thus
%$\hat{Y}(z)=Y(z)+\nu(z)$.  The measurement noise $\nu(k)$ is enough for
%represent both quantisation noises ($\nu_{1}$ and $\nu_{2}$) what can be
%obtained reorganizing the equations~\ref{eq:outputfunctions}:
%%
%\begin{equation}
%\label{eq:outputfunctions2}
%\hat{Y}(z)=H_{1}R(z)+H_{1}\nu_{1}(z)+H_{2}(z)\nu_{2}(z),
%\end{equation}
%%
%where $H_{1}R(z)$ is exactly equal to $Y(z)$ without quantisation noise effects, thus:
%%
%\begin{equation}
%\nu(z)=H_{1}\nu_{1}(z)+H_{2}(z)\nu_{2}(z).
%\end{equation}

@@

%
In Figure~\ref{fig:digsystem1}, two sources of uncertainties are
illustrated: parametric uncertainties related to modelling errors, and
uncertainties introduced by quantisations in sampling and holding process,
i.e., ADC and DAC conversions.  Assuming linearity and independence between
these sources, the uncertain model may be rewritten as
%
\begin{equation}
\label{eq:complete_unc_model}
\vec{\hat{G}}=\vec{G}+\Delta_p \vec{G}+\Delta_q \vec{G},
\end{equation}
%
where $\Delta_p \vec{G}$ is the term that corresponds to the parametric
uncertainties and $\Delta_q \vec{G}$ is the term that reflects the
quantisation effects on the plant model.

@@

Since the controller is implemented using a FWL, $C(z)$ also suffers
disturbances from the FWL effects, with roundoffs in coefficients that may
change the closed-loop poles and zeros position, and consequently affect its
stability, as argued in~\cite{Bessa16}.

Let $\hat{C}(z)$ be the transfer function of the digital controller
represented using this FWL with integer size $I$ and fractional size~$F$. 
We shall denote the FWL domain of coefficients as $\mathbb{R}\langle I,F
\rangle$
%
%Since the controller is implemented in the $\mathbb{R}\langle I,F \rangle$ 
%domain, {\it i.e.} a Fixed-Point Finite Word Length with integer size $I$ and fractional
%size $F$, we shall denote this domain as $\mathbb{R}\langle I,F \rangle$ 
%
and assume that there exists a function
$FWL[\cdot]:\mathcal{P}^{n}\rightarrow \mathcal{P}^{n}\langle I,F \rangle$,
which applies the FWL effects to a polynomial.  $\mathcal{P}^{n}$ is the
space of polynomials of $n$-th order and $\mathcal{P}^{n}\langle I,F
\rangle$ is is the space of polynomials with coefficients in
$\mathbb{R}\langle I,F \rangle$.
%
%, \red{[explain index n! why do you need to work on a Euclidean domain?]}. 
%
Thus, the perturbed controller model $\hat{C}(z)$ may be obtained from the
original model $C(z)=\frac{N_{C}(z)}{D_{C}(z)}$ using the following
expression:
%
\begin{equation}
\hat{C}(z)=\frac{FWL[N_{C}(z)]}{FWL[D_{C}(z)]}.
\end{equation}
%
In the case of a synthesised controller, because the synthesis is performed
using FWL, $\hat{C}(z) \equiv C(z)$.
